---
title: "Chapter 1. Data Visualization"
author: "Joe Paris"
date: "1/11/2023"
format: 
  html:
    fig-width: 8
    fig-height: 4
    code-fold: true
---

## Introduction

```{r}
pacman::p_load("tidyverse")

# tidyverse_update()
```

```{r}
pacman::p_load("palmerpenguins")
pacman::p_load(ggthemes)
```

```{r}
palmerpenguins::penguins
```

```{r}
glimpse(penguins)
```

```{r}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = aes(color = species, shape = species)) +
  geom_smooth(method = "lm") +
  labs(
    title = "Body mass and flipper length",
    subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
    x = "Flipper length (mm)",
    y = "Body mass (g)",
    color = "Species",
    shape = "Species"
  )
```

### Exercises

1.  How many rows are in `penguins`? How many columns?

    344 rows (or observations) and 8 columns (or variables).

2.  What does the `bill_depth_mm` variable in the `penguins` data frame describe? Read the help for `?penguins` to find out.

    "a number denoting bill depth (millimeters)"

3.  Make a scatterplot of `bill_depth_mm` versus `bill_length_mm.` That is, make a scatterplot with `bill_depth_mm` on the y-axis and `bill_length_mm` on the x-axis. Describe the relationship between these two variables.

    ```{r}
    ggplot(
      data = penguins,
      mapping = aes(
    x = bill_depth_mm,
    y = bill_length_mm,
    color = species,
    shape = species
      )
    ) +
      geom_point() +
      geom_smooth(method = "lm") +
      labs(
    title = "Bill depth and length",
    subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
    x = "Bill depth (mm)",
    y = "Bill length (mm)",
    color = "Species",
    shape = "Species"
      )
    ```

    There appears to be a linear relationship between bill depth and length. ("There is a positive, linear, and somewhat strong association between bill depth and length.")

4.  What happens if you make a scatterplot of `species` versus `bill_depth_mm`? What might be a better choice of geom?

    ```{r}
    ggplot(
      data = penguins,
      mapping = aes(
    x = bill_depth_mm,
    y = species,
    color = species,
    shape = species
      )
    ) +
      geom_point() +
      geom_smooth(method = "lm") +
      labs(
    title = "Species and bill depth",
    x = "Bill depth (mm)",
    y = "Species",
    color = "Species",
    shape = "Species"
      )
    ```

    Species is a categorical variable and a scatterplot of a categorical variable is not that useful as it’s difficult to use it to describe the distribution of bill depth across species.

    A better choice of geom might be a bar chart showing average bill length per species.

5.  Why does the following give an error, and how would you fix it?

    ``` r
    ggplot(data = penguins) + 
      geom_point()
    ```

    `geom_point` requires both `x` and `y` aesthetic values.

6.  What does the `na.rm` argument do in `geom_point()`? What is the default value of the argument? Create a scatterplot where you successfully use this argument set to `TRUE`.

    |         |                                                                                                                 |
    |-------------------|-----------------------------------------------------|
    | `na.rm` | If FALSE, the default, missing values are removed with a warning. If TRUE, missing values are silently removed. |

    ```{r}
    ggplot(
      data = penguins,
      mapping = aes(x = flipper_length_mm, y = body_mass_g)
    ) +
      geom_point(
    mapping = aes(color = species, shape = species),
    na.rm = TRUE
      ) +
      labs(
    title = "Body mass and flipper length",
    subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
    x = "Flipper length (mm)",
    y = "Body mass (g)",
    color = "Species",
    shape = "Species"
      )
    ```

7.  Add the following caption to the plot you made in the previous exercise: “Data come from the palmerpenguins package.” Hint: Take a look at the documentation for `labs()`.

    ```{r}
    ggplot(
      data = penguins,
      mapping = aes(x = flipper_length_mm, y = body_mass_g)
    ) +
      geom_point(
    mapping = aes(color = species, shape = species),
    na.rm = TRUE
      ) +
      labs(
    title = "Body mass and flipper length",
    subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
    caption = "Data come from the palmerpenguins package.",
    x = "Flipper length (mm)",
    y = "Body mass (g)",
    color = "Species",
    shape = "Species"
      )
    ```

8.  Re-create the following visualization. What aesthetic should `bill_depth_mm` be mapped to? And should it be mapped at the global level or at the geom level?

    ```{r}
    ggplot(
      data = penguins,
      mapping = aes(x = flipper_length_mm, y = body_mass_g)
    ) +
      geom_point(
    mapping = aes(color = bill_depth_mm, shape = species),
      ) +
      geom_smooth() +
      labs(
        x = "Flipper length (mm)",
        y = "Body mass (g)",
        color = "Bill depth (mm)",
        shape = "Species"
      )
    ```

9.  Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.

    ```{r}
    ggplot(
      data = penguins,
      mapping = aes(
        x = flipper_length_mm,
        y = body_mass_g,
        color = island
      )
    ) +
      geom_point() +
      geom_smooth(se = FALSE)
    ```

    I used `?geom_smooth` to discover what the `se` parameter did.

10. Will these two graphs look different? Why/why not?

    ```{{r}}
    ggplot(
      data = penguins,
      mapping = aes(x = flipper_length_mm, y = body_mass_g)
    ) +
      geom_point() +
      geom_smooth()

    ggplot() +
      geom_point(
        data = penguins,
        mapping = aes(x = flipper_length_mm, y = body_mass_g)
      ) +
      geom_smooth(
        data = penguins,
        mapping = aes(x = flipper_length_mm, y = body_mass_g)
      )
    ```

    "The graphs will look the same as the first is using the default first two arguments."The two plots will look the same as in the first plot the aesthetic mappings are at the global level and passed down to both geoms, and in the second plot both geoms have the same aesthetic mappings, each defined at the local level."

## ggplot2 Calls

